---
title: "Deltas Notebook"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
  word_document: default
editor_options: 
  chunk_output_type: inline
---

Load packages
```{r load-packages}
library(tidyverse)
library(gridExtra)
library(maps)
library(ggthemes)

```

Let's import all of the data that Simon pulled:
```{r import-data}
## Original
# DeltasClean <- read_csv('../data/out/deltas_clean_v2.csv')
# DeltaLocations <- read_csv("../data/DeltaLocations.csv")

## With Added Deltas
DeltasClean <- read_csv("data/out/deltas_clean_96.csv")
DeltaLocations <- read_csv("data/DeltaLocations_96.csv")
```


As a reminder, for each of the 47 deltas there are measurements of Land & Water areas at Upstream, Downstream and 'Middle' locations on the delta. We first lump all the observations together, and look to see which Deltas have many observations:

```{r delta-counts}
#counts per delta
DeltaCounts <- count(DeltasClean, Delta)
DeltaCounts
```


In the heat map below, dark colors (and no color) represent data paucity (and data gaps). Deltas with light colors (e.g., the Parana, Nile, Ebro, Colorado, Brahmani) have lots of data, spread out through the months of the year.

Now, by each month.. where the colorbar represents the number of observations (n) for each month for a given delta:

```{r delta-obs-by-month}
DeltaObsPerMonth <- count(DeltasClean, Delta, month)

ggplot(DeltaObsPerMonth, aes(y = Delta, x = factor(month), fill=n)) + geom_tile() + 
  scale_x_discrete(limits = c(1:12), breaks = c(1:12)) +
  expand_limits(x = c(1,12)) + 
  scale_fill_gradient( trans = 'log' )
```


I'll remove/subset the deltas with sparse coverage (specifically, months with no coverage).... 
```{r remove-sparse-deltas}

# need 10 data points per month for NDSSI and NDVI
EnoughObsPerMonth <- DeltasClean %>% ungroup() %>%
  count(Delta, month, surface) %>% 
  group_by(surface) %>%
  filter( n >= 5)

#find deltas missing a given month of observations
DeltaMonthCounts <- EnoughObsPerMonth %>%
  ungroup() %>%
  count(Delta)

# need 12 months of water and land obs, so 24 mo total
EnoughMonths <- DeltaMonthCounts %>%
 filter( n == 24)

CompleteObsDeltas <- pull(EnoughMonths, Delta)

#remove them
DeltasCleaner <- DeltasClean %>%
  filter(Delta %in% CompleteObsDeltas)

#add the real dates in month date format
DeltasCleaner$date <- as.Date(paste(DeltasCleaner$year, DeltasCleaner$month, "01", sep="-"), "%Y-%m-%d")

#remove intermediate data
rm(CompleteObsDeltas, EnoughMonths, DeltaMonthCounts)

#EnoughMonths
```

and extract some metrics; specifically I will make a timeseries of NDVI and NDSSI for each delta using the mean value for each month. Added NSMI - JM

```{r delta_means_max_min, include = TRUE}
#take the mean NDVI, NDSSI, NSMI for each month, for each delta
DeltaMeans <- DeltasCleaner %>%
  group_by(Delta, month, surface) %>%
  summarize(MeanNDVI = mean(ndvi, na.rm = TRUE), 
            MeanNDSSI = mean(ndssi, na.rm = TRUE),
            MeanNSMI = mean(nsmi, na.rm = TRUE),
            MeanEVI = mean(evi, na.rm = TRUE))

#make a 11 column data frame with:
#delta, 
#max and min NDVI month, 
#NDSSI max and min month, 
#max and min values for both NDVI and NDSSI

#NDVI
DeltaMaxNDVI <- 
  DeltaMeans %>% 
  filter(surface == 'Land')  %>% 
  select (-c(MeanNDSSI, MeanNSMI, MeanEVI, surface)) %>% 
  group_by(Delta) %>% 
  slice(which.max(MeanNDVI)) %>% 
  rename(MaxMeanNDVImonth = month, MaxMeanNDVI = MeanNDVI)

DeltaMinNDVI <- 
  DeltaMeans %>% 
  filter(surface == 'Land')  %>% 
  select (-c(MeanNDSSI, MeanNSMI, MeanEVI, surface)) %>% 
  group_by(Delta) %>% 
  slice(which.min(MeanNDVI)) %>% 
  rename(MinMeanNDVImonth = month, MinMeanNDVI = MeanNDVI)

#NDSSI
DeltaMaxNDSSI <- 
  DeltaMeans %>% 
  filter(surface == 'Water')  %>% 
  select (-c(MeanNDVI, MeanNSMI, MeanEVI, surface)) %>% 
  group_by(Delta) %>% 
  slice(which.max(MeanNDSSI)) %>% 
  rename(MaxMeanNDSSImonth = month, MaxMeanNDSSI = MeanNDSSI)

DeltaMinNDSSI <- 
  DeltaMeans %>% 
  filter(surface == 'Water')  %>% 
  select (-c(MeanNDVI, MeanNSMI, MeanEVI, surface)) %>% 
  group_by(Delta) %>% 
  slice(which.min(MeanNDSSI)) %>% 
  rename(MinMeanNDSSImonth = month, MinMeanNDSSI = MeanNDSSI)

#NSMI
DeltaMaxNSMI <- 
  DeltaMeans %>% 
  filter(surface == 'Water') %>% 
  select(-c(MeanNDVI, MeanNDSSI, MeanEVI, surface)) %>% 
  group_by(Delta) %>% 
  slice(which.max(MeanNSMI)) %>% 
  rename(MaxMeanNSMImonth = month, MaxMeanNSMI = MeanNSMI)

DeltaMinNSMI <-
  DeltaMeans %>% 
  filter(surface == 'Water') %>% 
  select(-c(MeanNDVI, MeanNDSSI, MeanEVI, surface)) %>% 
  group_by(Delta) %>% 
  slice(which.min(MeanNSMI)) %>% 
  rename(MinMeanNSMImonth = month, MinMeanNSMI = MeanNSMI)

#EVI
DeltaMaxEVI <- 
  DeltaMeans %>% 
  filter(surface == 'Land')  %>% 
  select (-c(MeanNDSSI, MeanNSMI, MeanNDVI, surface)) %>% 
  group_by(Delta) %>% 
  slice(which.max(MeanEVI)) %>% 
  rename(MaxMeanEVImonth = month, MaxMeanEVI = MeanEVI)

DeltaMinEVI <- 
  DeltaMeans %>% 
  filter(surface == 'Land')  %>% 
  select (-c(MeanNDSSI, MeanNSMI, MeanNDVI, surface)) %>% 
  group_by(Delta) %>% 
  slice(which.min(MeanEVI)) %>% 
  rename(MinMeanEVImonth = month, MinMeanEVI = MeanEVI)
  

#join into 1 dataframe
DeltaMaxMin <- left_join(DeltaMaxNDVI, DeltaMaxNDSSI, by = 'Delta') %>% 
  left_join(.,DeltaMinNDVI, by = 'Delta') %>% 
  left_join(.,DeltaMinNDSSI, by = 'Delta') %>% 
  left_join(.,DeltaMaxNSMI, by = 'Delta') %>% 
  left_join(.,DeltaMinNSMI, by = 'Delta') %>% 
  left_join(.,DeltaMaxEVI, by = 'Delta') %>% 
  left_join(.,DeltaMinEVI, by = 'Delta')
#remove intermediate data
rm(DeltaMaxNDVI, DeltaMaxNDSSI, DeltaMinNDSSI,DeltaMinNDVI, DeltaMaxNSMI, DeltaMinNSMI, DeltaMaxEVI, DeltaMinEVI)

```


And now we can look at the phase shifts between these two time series (the timeseries of mean NDVI and mean NDSSI). Here are the phase shifts (in month) for each delta:
```{r }
#compare offset
DeltaMaxMin <- mutate(DeltaMaxMin, 
                      MinOffset = if_else(abs(MinMeanNDVImonth - MinMeanNDSSImonth) > 6, 
                                          12 - abs(MinMeanNDVImonth - MinMeanNDSSImonth),
                                          abs(MinMeanNDVImonth - MinMeanNDSSImonth)),
                      MaxOffset = if_else(abs(MaxMeanNDVImonth - MaxMeanNSMImonth) > 6, 
                                         12 - abs(MaxMeanNDVImonth - MaxMeanNSMImonth),
                                          abs(MaxMeanNDVImonth - MaxMeanNSMImonth)),
                      MaxOffsetEVI = if_else(abs(MaxMeanEVImonth - MaxMeanNSMImonth) > 6, 
                                         12 - abs(MaxMeanEVImonth - MaxMeanNSMImonth),
                                          abs(MaxMeanEVImonth - MaxMeanNSMImonth)),
                      MaxMinOffset = if_else(abs(MaxMeanNDVImonth - MinMeanNDSSImonth) > 6, 
                                         12 - abs(MaxMeanNDVImonth - MinMeanNDSSImonth),
                                          abs(MaxMeanNDVImonth - MinMeanNDSSImonth)),
                      OffsetDiff = abs(MaxOffset - MinOffset),
                      rangeNDVI = (MaxMeanNDVI - MinMeanNDVI), 
                      rangeNDSSI = (MaxMeanNDSSI - MinMeanNDSSI),
                      rangeNSMI = (MaxMeanNSMI - MinMeanNSMI),
                      rangeEVI = (MaxMeanEVI - MinMeanEVI),
                      halfPeriodNDVI = if_else(abs(MaxMeanNDVImonth - MinMeanNDVImonth) > 6, 
                                          12 - abs(MaxMeanNDVImonth - MinMeanNDVImonth),
                                          abs(MaxMeanNDVImonth - MinMeanNDVImonth)),
                      halfPeriodNDSSI = if_else(abs(MaxMeanNDSSImonth - MinMeanNDSSImonth) > 6, 
                                          12 - abs(MaxMeanNDSSImonth - MinMeanNDSSImonth),
                                          abs(MaxMeanNDSSImonth - MinMeanNDSSImonth)), )

# DeltaMaxMin <- 
#   DeltaMaxMin   %>%
#   select (c(Delta, MinOffset, MaxOffset, OffsetDiff, rangeNDVI, rangeNDSSI,MaxMeanNDSSI,MinMeanNDSSI,MaxMeanNDVI,MinMeanNDVI))
```


```{r }
DeltaMaxMin

ggplot(DeltaMaxMin, aes(y = Delta, x = MaxOffset)) + geom_point() + 
  scale_x_discrete(limits = c(1:6), breaks = c(1:6)) +
  expand_limits(x = c(0,6))  + 
  ggtitle("MaxOffset")

ggplot(DeltaMaxMin, aes(y = Delta, x = MinOffset)) + geom_point() + 
  scale_x_discrete(limits = c(1:6), breaks = c(1:6)) +
  expand_limits(x = c(0,6))  + 
  ggtitle("MinOffset")

ggplot(DeltaMaxMin, aes(y = Delta, x = OffsetDiff)) + geom_point() + 
  scale_x_discrete(limits = c(1:6), breaks = c(1:6)) +
  expand_limits(x = c(0,6))  + 
  ggtitle("Offset Difference")

```

Now let's examine the histograms of all 31 deltas... The months with the greatest mean NDVI, months with gretaest mean NDSSI, the monthly offset, and the skew of the NDSSI and NDVI timeseries.

```{r}
ggplot(DeltaMaxMin, aes(x = MaxMeanNDVImonth)) + 
  geom_dotplot(binwidth = 1,dotsize = 0.5) + 
  scale_y_continuous(NULL, breaks = NULL) + 
  scale_x_discrete(limits = c(1:12), breaks = c(1:12)) + 
  labs(x = "Month") +
  ggtitle("Month of maximum mean NDVI")

ggplot(DeltaMaxMin, aes(x = MaxMeanNDSSImonth)) + 
  geom_dotplot(binwidth = 1,dotsize = 0.5) + 
  scale_y_continuous(NULL, breaks = NULL) + 
  scale_x_discrete(limits = c(1:12), breaks = c(1:12)) + 
  labs(x = "Month") +
  ggtitle("Month of maximum mean NDSSI")

ggplot(DeltaMaxMin, aes(x = MaxOffset)) + 
  geom_dotplot(binwidth = 1,dotsize = 0.25) + 
  scale_y_continuous(NULL, breaks = NULL) + 
  scale_x_discrete(limits = c(0:6), breaks = c(0:6)) + 
  labs(x = "Months") +
  ggtitle("Months Offset between NDVI and NDSSI")

ggplot(DeltaMaxMin, aes(x = halfPeriodNDVI)) + 
  geom_dotplot(binwidth = 1,dotsize = 0.25) + 
  scale_y_continuous(NULL, breaks = NULL) + 
  scale_x_discrete(limits = c(0:6), breaks = c(0:6)) + 
  labs(x = "Months") +
  ggtitle("half period NDVI ")

ggplot(DeltaMaxMin, aes(x = halfPeriodNDSSI)) + 
  geom_dotplot(binwidth = 1,dotsize = 0.25) + 
  scale_y_continuous(NULL, breaks = NULL) + 
  scale_x_discrete(limits = c(0:6), breaks = c(0:6)) + 
  labs(x = "Months") +
  ggtitle("half period NDSSI")

```

Ok, so the idea is that peak NDSSI is more effective if it occurs at moderate NDVI, so let's look at the NDVI value for the months with peak NDSSI. 
```{r}
#extract NDVI value for each delta a the month of max NDSSI value

DeltaNDVIatMaxNDSSI <- DeltaMaxMin %>%
  select(Delta,MaxMeanNDSSImonth)

DeltaMeansToJoin <- DeltaMeans %>%
  filter(surface == 'Land')

DeltaNDVIatMaxNDSSI <- left_join(DeltaNDVIatMaxNDSSI, DeltaMeansToJoin, 
                                 by = c('Delta', 'MaxMeanNDSSImonth' ='month'))
 
DeltaNDVIatMaxNDSSI <- DeltaNDVIatMaxNDSSI %>%
  select (-c(surface, MeanNDSSI))

DeltaNDVIatMaxNDSSI

M <- DeltaNDVIatMaxNDSSI %>%
  ggplot( aes(x=MeanNDVI)) +
    geom_histogram( color="#e9ecef", alpha=0.6)

M
   
ggplot(DeltaNDVIatMaxNDSSI, aes(x = MeanNDVI)) + 
  geom_dotplot(binwidth = 0.1, dotsize = 0.25) + 
  scale_y_continuous(NULL, breaks = NULL)  + 
  xlim(0,1) +
  labs(x = "NDVI") +
  ggtitle("NDVI at month of maximum mean NDSSI")
```

And what if NDVI peak was one month sooner:
```{r}
#extract NDVI value for each delta at one month earlier than max NDSSI value

DeltaNDVIatEarlyNDSSI <- DeltaMaxMin %>%
  select(Delta,MaxMeanNDSSImonth)

DeltaNDVIatEarlyNDSSI <- DeltaNDVIatEarlyNDSSI %>% 
  mutate(EarlyNDSSI = if_else(MaxMeanNDSSImonth == 1, 
                              12,
                              MaxMeanNDSSImonth-1)
         )

DeltaNDVIatEarlyNDSSI <- DeltaNDVIatEarlyNDSSI %>%
  select (-c(MaxMeanNDSSImonth))

DeltaMeansToJoin <- DeltaMeans %>%
  filter(surface == 'Land')

DeltaNDVIatEarlyNDSSI <- left_join(DeltaNDVIatEarlyNDSSI, DeltaMeansToJoin, 
                                 by = c('Delta', 'EarlyNDSSI' ='month'))
 
DeltaNDVIatEarlyNDSSI <- DeltaNDVIatEarlyNDSSI %>%
  select (-c(surface, MeanNDSSI))

DeltaNDVIatEarlyNDSSI

n <- DeltaNDVIatEarlyNDSSI %>%
  ggplot( aes(x=MeanNDVI)) +
    geom_histogram( color="#e9ecef", alpha=0.6)

n

ggplot(DeltaNDVIatEarlyNDSSI, aes(x = MeanNDVI)) + 
  geom_dotplot(binwidth = 0.1, dotsize = 0.25) + 
  scale_y_continuous(NULL, breaks = NULL)  + 
  xlim(0,1) +
  labs(x = "NDVI") +
  ggtitle("NDVI at 1 month earlier than maximum mean NDSSI")
```


How about the NDVI for months with the lowest NDSSI

```{r}
#extract NDVI value for each delta a the month of min NDSSI value

DeltaNDVIatMinNDSSI <- DeltaMaxMin %>%
  select(Delta,MinMeanNDSSImonth)

DeltaMeansToJoin <- DeltaMeans %>%
  filter(surface == 'Land')

DeltaNDVIatMinNDSSI <- left_join(DeltaNDVIatMinNDSSI, DeltaMeansToJoin, 
                                 by = c('Delta', 'MinMeanNDSSImonth' ='month'))
 
DeltaNDVIatMinNDSSI <- DeltaNDVIatMinNDSSI %>%
  select (-c(surface, MeanNDSSI))

DeltaNDVIatMinNDSSI

p <- DeltaNDVIatMinNDSSI %>%
  ggplot( aes(x=MeanNDVI)) +
    geom_histogram( color="#e9ecef", alpha=0.6)

p

ggplot(DeltaNDVIatMinNDSSI, aes(x = MeanNDVI)) + 
  geom_dotplot(binwidth = 0.1, dotsize = 0.25) + 
  scale_y_continuous(NULL, breaks = NULL)  + 
  xlim(0,1) +
  labs(x = "NDVI") +
  ggtitle("NDVI at month of min mean NDSSI")
```


Just to explore the data a bit, here are the phase shifts/offsets against other measured parameters for each delta. The range, max and mean of NDVI and NDSSI is calculated from the timeseries, so it is really the max, min, and range of the monthly means (i.e., the maximum of the means, the minimum of the means, and the range of the mean). Offset is measured in months.

```{r}
slice1 <- ggplot(DeltaMaxMin, aes(y = rangeNDVI, x = MaxOffset)) + geom_point() 
# + scale_x_discrete(limits = c(1:6), breaks = c(1:6)) + expand_limits(x = c(1,6)) 

slice2 <- ggplot(DeltaMaxMin, aes(y = rangeNDVI, x = rangeNDSSI)) + geom_point() 
slice3 <- ggplot(DeltaMaxMin, aes(y = rangeNDSSI, x = MaxOffset)) + geom_point() 

slice4 <- ggplot(DeltaMaxMin, aes(y = MaxMeanNDVI, x = rangeNDVI)) + geom_point() 
slice5 <- ggplot(DeltaMaxMin, aes(y = MaxMeanNDVI, x = rangeNDSSI)) + geom_point() 
slice6 <- ggplot(DeltaMaxMin, aes(y = MaxMeanNDVI, x = MaxOffset)) + geom_point() 

slice7 <- ggplot(DeltaMaxMin, aes(y = MaxMeanNDSSI, x = MaxMeanNDVI)) + geom_point() 
slice8 <- ggplot(DeltaMaxMin, aes(y = MaxMeanNDSSI, x = rangeNDSSI)) + geom_point() 
slice9 <- ggplot(DeltaMaxMin, aes(y = MaxMeanNDSSI, x = MaxOffset)) + geom_point() 
slice10 <- ggplot(DeltaMaxMin, aes(y = MaxMeanNDSSI, x = rangeNDVI)) + geom_point() 

grid.arrange(slice1, slice2, slice3, slice4, slice5, slice6, ncol=3)
grid.arrange(slice7, slice8, slice9, slice10, ncol=2)

#remove those panels
rm(slice1, slice2, slice3, slice4, slice5, slice6,slice7, slice8, slice9, slice10)

```


Join Latitude and longitude data

```{r add-location}
DeltaDatawLocations <- left_join(DeltaMaxMin, DeltaLocations, by = c("Delta" = "Deltas"))

DeltaDatawLocations <- DeltaDatawLocations %>%
  mutate(Absolute_Latitude= abs(Lat))
```

plot params vs lat
```{r}
loc1 <- ggplot(DeltaDatawLocations, aes(y = Absolute_Latitude, x = MaxOffset)) + geom_point() 
loc2 <- ggplot(DeltaDatawLocations, aes(y = Absolute_Latitude, x = rangeNDSSI)) + geom_point() 
loc3 <- ggplot(DeltaDatawLocations, aes(y = Absolute_Latitude, x = rangeNDVI)) + geom_point() + theme_few()
loc4 <- ggplot(DeltaDatawLocations, aes(y = Absolute_Latitude, x = MaxMeanNDVI)) + geom_point() 
loc5 <- ggplot(DeltaDatawLocations, aes(y = Absolute_Latitude, x = MaxMeanNDSSI)) + geom_point() 

loc6 <- ggplot(DeltaDatawLocations, aes(y = Absolute_Latitude, x = MaxOffsetNSMI)) + geom_point() + theme_few()
loc7 <- ggplot(DeltaDatawLocations, aes(y = Absolute_Latitude, x = rangeNSMI)) + geom_point() + theme_few()
loc8 <- ggplot(DeltaDatawLocations, aes(y = Absolute_Latitude, x = MaxMinOffset)) + geom_point() 

# grid.arrange(loc1, loc2, loc3, loc4, loc5, ncol=2)

loc1
#ggsave("loc1.pdf", width = 4, height = 4)
loc2
#ggsave("loc2.pdf", width = 4, height = 4)
loc3
ggsave("NDVI.png", dpi = 300, height = 5, width = 5, unit = 'in')

loc6
ggsave("OffsetNSMI.png", dpi = 300, height = 5, width = 5, unit = 'in')
loc7
ggsave("NSMI.png", dpi = 300, height = 5, width = 5, unit = 'in')

loc8

#remove those panels
rm(loc1, loc2, loc3, loc4, loc5)
```

```{r}
#find the linear model 
DeltaOffset_lm <- lm( Absolute_Latitude ~ MaxOffset, data = DeltaDatawLocations) 

summary(DeltaOffset_lm)

ggplot(DeltaDatawLocations, aes(x = Absolute_Latitude, y = MaxOffset)) + 
  geom_point() +
  geom_smooth(mapping = aes(x = Absolute_Latitude, y = MaxOffset, ), method=lm ) 



```


Now for some maps of the data maps:

```{r}
world <- ggplot() +
  borders("world", colour = "gray85", fill = "gray80") +
  theme_map() 

DeltaOffsetMap <- world +
  geom_point(aes(x = Lon, y = Lat, color = MaxOffsetNSMI),
             data = DeltaDatawLocations, 
             size = 5) + scale_color_gradient( high = "red", low  = "yellow")

DeltaOffsetMap
ggsave("DeltaOffsetMap.png")

```


```{r}
world <- ggplot() +
  borders("world", colour = "gray85", fill = "gray80") +
  theme_map() 

DeltaNDVIrangeMap <- world +
  geom_point(aes(x = Lon, y = Lat, color = rangeNDVI),
             data = DeltaDatawLocations,
             size = 5) + scale_color_gradient( high = "red", low  = "yellow") + 
  ggtitle("NDVI range")


DeltaNDSSIrangeMap  <- world +
  geom_point(aes(x = Lon, y = Lat, color = rangeNDSSI),
             data = DeltaDatawLocations, 
             size = 5) + scale_color_gradient( high = "red", low = "yellow") + ggtitle("NDSSI range") 

DeltaNSMIrangeMap <- world +
  geom_point(aes(x = Lon, y = Lat, color = rangeNSMI),
             data = DeltaDatawLocations, 
             size = 5) + scale_color_gradient( high = "red", low = "yellow") + ggtitle("NSMI range") 


DeltaNDVIrangeMap 
ggsave("DeltaNDVIrangeMap.png")
DeltaNDSSIrangeMap
#ggsave("DeltaNDSSIrangeMap.pdf", width = 6, height = 4)
DeltaNSMIrangeMap
ggsave("DeltaNSMIrangeMap.png")

```

Let's look at some of the timeseries 
To quantify the water, we use NDSSI. to quantify land, we use NDVI.

Changed to NSMI in function below - JM

First here is the function to make the plots:

```{r timeSeriesPlots}
DeltaPlotter <- function(DeltaName) {
  #Counts each month
  numVeg <- DeltasCleaner %>%
    select(Delta, surface, month, ndvi) %>%
    filter(Delta == DeltaName & 
             surface == "Land" & !is.na(ndvi)) %>%
    group_by(month) %>%
    summarize(n = n())
  
  numSed <- DeltasCleaner %>%
    select(Delta, surface, month, nsmi) %>%
    filter(Delta == DeltaName &
             surface == "Water" & !is.na(nsmi)) %>%
    group_by(month) %>%
    summarize(n = n())
  
  #Highlight the Maximum and Minimum Month for each delta, NDVI and NDSSI
  
  #LAND
  Veg <-
    ggplot(data = filter(DeltasCleaner, Delta == DeltaName &
                           surface == "Land")) +
    geom_boxplot(aes(x = month, y = ndvi, group = month)) +
    scale_x_discrete(limits = c(1:12), breaks = c(1:12)) +
    expand_limits(x = c(1, 12)) +
    ggtitle(DeltaName) +
    #geom_text(data = numVeg, aes(y = 1.05, x = month, label = n)) +
    geom_boxplot(
      data = filter(
        DeltasCleaner,
        Delta == DeltaName &
          surface == "Land" & month == DeltaMaxMin$MaxMeanNDVImonth[DeltaMaxMin$Delta == DeltaName] 
      ),
      aes(x = month, y = ndvi, group = month),
      fill = "green"
    ) +
    geom_boxplot(
      data = filter(
        DeltasCleaner,
        Delta == DeltaName & 
          surface == "Land" & month == DeltaMaxMin$MinMeanNDVImonth[DeltaMaxMin$Delta ==DeltaName]
      ),
      aes(x = month, y = ndvi, group = month),
      fill = "blue"
    )
  
  
  Sed <-
    ggplot(data = filter(DeltasCleaner, Delta == DeltaName &
                           surface == "Water")) +
    geom_boxplot(aes(x = month, y = nsmi, group = month)) +
    scale_x_discrete(limits = c(1:12), breaks = c(1:12)) +
    expand_limits(x = c(1, 12)) +
    #geom_text(data = numSed, aes(y = 1.05, x = month, label = n)) +
    geom_boxplot(
      data = filter(
        DeltasCleaner,
        Delta == DeltaName &
          surface == "Water" & month == DeltaMaxMin$MaxMeanNSMImonth[DeltaMaxMin$Delta == DeltaName]
      ),
      aes(x = month, y = nsmi, group = month),
      fill = "green"
    ) +
    geom_boxplot(
      data = filter(
        DeltasCleaner,
        Delta == DeltaName &
          surface == "Water" & month == DeltaMaxMin$MinMeanNSMImonth[DeltaMaxMin$Delta == DeltaName]
      ),
      aes(x = month, y = nsmi, group = month),
      fill = "blue"
    )
  
  return(grid.arrange(Veg, Sed, nrow = 2))
}
```


Here is are some examples:

* The peaks in both timeseries shift around depending on the delta:
   + look at the correlation in the Orinoco nad Senegal
   + The anticorrelation in the Parana and Ebro,
   + The slight phase shift in the Magdalena.

```{r}
Magdalena <- DeltaPlotter("Magdalena")
ggsave("Magdalena.png", plot = Magdalena)
Nile <- DeltaPlotter("Nile")
ggsave("Nile.png", plot = Nile)
Mississippi <- DeltaPlotter("Mississippi")
ggsave("Miss.png", plot = Mississippi)
Columbia <- DeltaPlotter("Columbia")
ggsave("Columbia.png", plot = Columbia)

# 
# DeltaPlotter("Parana")
# DeltaPlotter("Magdalena")
# DeltaPlotter("Ebro")
# DeltaPlotter("Nile")
# DeltaPlotter("Senegal")
# DeltaPlotter("Orinoco")
# DeltaPlotter("Godavari")
# DeltaPlotter("Krishna")
# DeltaPlotter("Kampar")
# DeltaPlotter("Parnaiba")
# DeltaPlotter("Fly")
# DeltaPlotter("Mekong")
# DeltaPlotter("Sepik")
# DeltaPlotter("RioG")
# DeltaPlotter("Mississippi")
# DeltaPlotter("Mobile")
```


Now for some work with GRDC discharge data:
```{r GRDC}
#import the data (monthly means for 21 stations)
DeltasGRDC  <- read_csv("data/GRDCstations.csv")

#rename the months by numbers and tidy the GRDC data
DeltasDischarge <- DeltasGRDC %>%
  rename("1" = January, "2"= February, "3"= March, "4"= April,
         "5"=May, "6"=June, "7"=July, "8"= August, "9" = September, "10"=October, 
         "11"=November, "12"=December) %>%
  select(Delta, "1" , "2" , "3", "4","5", "6", "7", "8", "9", "10", "11", "12") %>%
  pivot_longer(-Delta, names_to = "month", values_to = "discharge")

#find max and min GRDC month for each delta
DeltaMaxDischarge <- 
  DeltasDischarge %>% 
  group_by(Delta) %>% 
  slice(which.max(discharge)) %>% 
  rename(MaxDischargeMonth = month, MaxDischarge = discharge) %>% 
  ungroup()

DeltaMinDischarge <-
  DeltasDischarge %>% 
  group_by(Delta) %>% 
  slice(which.min(discharge)) %>% 
  rename(MinDischargeMonth = month, MinDischarge = discharge) %>% 
  ungroup()

DeltaMaxDischarge$MaxDischargeMonth = as.numeric(DeltaMaxDischarge$MaxDischargeMonth)
DeltaMinDischarge$MinDischargeMonth = as.numeric(DeltaMinDischarge$MinDischargeMonth)

#join with other delta data
DeltawLocGRDC <- right_join(DeltaDatawLocations, DeltaMaxDischarge, by = 'Delta')
DeltawLocGRDC <- right_join(DeltawLocGRDC, DeltaMinDischarge, by = "Delta")

#calculate max offset, range
DeltawLocGRDC <- DeltawLocGRDC %>%
  mutate(DisMaxOff = if_else(abs(MaxDischargeMonth - MaxMeanNSMImonth) > 6,
                            12 - abs(MaxDischargeMonth - MaxMeanNSMImonth),
                            abs(MaxDischargeMonth - MaxMeanNSMImonth))) %>%
  mutate(DisMinOff = if_else(abs(MinDischargeMonth - MinMeanNSMImonth) > 6,
                            12 - abs(MinDischargeMonth - MinMeanNSMImonth),
                            abs(MinDischargeMonth - MinMeanNSMImonth))) %>%  
  mutate(DisNDSSIOff = if_else(abs(MaxDischargeMonth - MaxMeanNDSSImonth) > 6,
                            12 - abs(MaxDischargeMonth - MaxMeanNDSSImonth),
                            abs(MaxDischargeMonth - MaxMeanNDSSImonth))) %>% 
  mutate(DisRange = MaxDischarge - MinDischarge)
  

```
Plot some delta figures

```{r GRDCFigs}
#Some histograms
ggplot(DeltawLocGRDC, aes(x = DisMaxOff)) + geom_histogram(binwidth = 1, color="black", fill="white")
ggplot(DeltawLocGRDC, aes(x = DisNDSSIOff)) + geom_histogram(binwidth = 1, color="black", fill="white")
ggplot(DeltawLocGRDC, aes(x = DisMinOff)) + geom_histogram(binwidth = 1, color="black", fill="white")

# #plot mean of monthly means against NDSSI
# ggplot(DeltawLocGRDC, aes(y = Range_Discharge, x = rangeNDSSI)) + geom_point() + scale_y_continuous(trans='log10')
# 
# #ggsave("GRDCNDSSI.pdf", width = 6, height = 4)
# 
# ggplot(DeltawLocGRDC, aes(y = Range_Discharge, x = MaxMeanNDSSI)) + geom_point() + scale_y_continuous(trans='log10')
# 
# #Compare offset with NDSSI (deltamaxmin$maxmeanNDSSImonth)
# ggplot(DeltaMaxMinDischarge, aes(y = Delta, x = DissOff)) + geom_point() + 
#   scale_x_discrete(limits = c(1:6), breaks = c(1:6)) +
#   expand_limits(x = c(0,6))  + 
#   ggtitle("DisOffset")
```

Look at GRDC data by latitude:

```{r}
#join lat data
DeltaDatawLocations <- left_join(DeltaMaxMinDischarge, DeltaLocations, by = c("Delta" = "Deltas"))

DeltaDatawLocations <- DeltaDatawLocations %>%
  mutate(Absolute_Latitude= abs(Lat))

# plot offset on graph by lat
ggplot(DeltaDatawLocations, aes(x = Absolute_Latitude, y = DissOff)) + geom_point() +
  scale_color_gradient(low = "yellow", high = "red", na.value = NA) 
  #+ geom_smooth(mapping = aes(x = Absolute_Latitude, y = DissOff, ), method=lm ) 

#ggsave("DisOffset.pdf", width = 6, height = 4)
```

And now on a map:
```{r}
#plot offset on map
DeltaDisOffsetMap <- world +
  geom_point(aes(x = Lon, y = Lat, color = DissOff),
             data = DeltaDatawLocations, 
             size = 5) + scale_color_gradient( high = "red", low  = "yellow") +
  ggtitle("Offset Between GRDC discharge peak and NDSSI peak in water")

DeltaDisOffsetMap
#ggsave("DeltaOffsetMap.pdf", width = 6, height = 4)

```

Calculate ASI between NDVI and NDSSI/NSMI
```{r calculate_ASI, warning=FALSE}
#Calculate ASI from Feng et al 2019, eq 1-3

#First get monthly mean NDVI, NSMI, NDSSI for each delta
DeltaMeansNDVI <- DeltaMeans %>% 
  filter(surface == "Land") %>% 
  select(Delta,month,MeanNDVI)

DeltaMeansNDSSI <- DeltaMeans %>% 
  filter(surface == "Water") %>% 
  select(Delta,month,MeanNDSSI)

DeltaMeansNSMI <- DeltaMeans %>% 
  filter(surface == "Water") %>% 
  select(Delta, month, MeanNSMI)

#Combine into one data frame
DeltaMeansASI <- left_join(DeltaMeansNDVI,DeltaMeansNDSSI, by = c("Delta", "month")) %>%
  left_join(.,DeltaMeansNSMI, by = c("Delta", "month")) %>% 
  mutate(mNDVI = MeanNDVI + 1, mNDSSI = MeanNDSSI + 1, mNSMI = MeanNSMI + 1) %>% #Transform both indices so non-negative
  select(Delta,month,mNDVI,mNDSSI,mNSMI)

#Now calculate Jansen-Shannon distance of observations
DeltaMeansASI <- DeltaMeansASI %>% #Requires more than one group/ungroup by Delta to get order correct
  group_by(Delta) %>% 
  mutate(sumNDVI = sum(mNDVI), sumNDSSI = sum(mNDSSI), sumNSMI = sum(mNSMI)) %>% #Sums of variables for pmf
  ungroup() %>% 
  mutate(pNDVI = mNDVI/sumNDVI, pNDSSI = mNDSSI/sumNDSSI, pNSMI = mNSMI/sumNSMI, #pmfs, mean pmf, information
         pM1 = 0.5 * (pNDVI + pNDSSI), pM2 = 0.5 * (pNDVI + pNSMI), #M1 is NDVI and NDSSI, M2 is NDVI and NSMI
         iNDVI1 = pNDVI * log2(pNDVI/pM1), iNDSSI1 = pNDSSI * log2(pNDSSI/pM1),
         iNDVI2 = pNDVI * log2(pNDVI/pM2), iNSMI2 = pNSMI * log2(pNSMI/pM2)) %>% 
  group_by(Delta) %>%
  mutate(D1ndvi = sum(iNDVI1), D1ndssi = sum(iNDSSI1),
         D2ndvi = sum(iNDVI2), D2nsmi = sum(iNSMI2),) %>%  #Calculate entropies for year
  ungroup()

JS <- DeltaMeansASI %>% 
  filter(.,DeltaMeansASI$month == 1) %>%
  mutate(JSobs1 = sqrt(0.5*D1ndvi + 0.5*D1ndssi),
         JSobs2 = sqrt(0.5*D2ndvi + 0.5*D2nsmi)) %>% 
  select(Delta,JSobs1,JSobs2) #JSobs1 is with NDSSI, JSobs2 is with NSMI

#Some functions to help with minJS calc
roll <- function( x , n ){ #function to help offset
  if( n == 0 )
    return( x )
  c( tail(x,n) , head(x,-n) )
}

calcMinJS <- function(x,y){ #Calculate JSmin from two vectors of monthly values NEED TO VECTORIZE
  JSall <- data.frame(offset = 0:11,JSw = 0)
  for (w in JSall$offset){
    temp <- data.frame(v1 = x, v2 = roll(y,w)) 
    JSall$JSw[w+1] <- temp %>% 
      mutate(pv1 = v1/sum(v1), pv2 = v2/sum(v2),
             M = 0.5*(pv1+pv2),
             dv1 = sum(pv1*log2(pv1/M)), dv2 = sum(pv2*log2(pv2/M)),
             JStemp = sqrt(0.5*dv1 + 0.5*dv2)) %>% 
      pull(JStemp[1])
  }
  min(JSall$JSw)
}

#Add JS min to JS
JS$JSmin1 <- DeltaMeansASI %>% 
  group_by(Delta) %>% 
  mutate(JSmin1 = calcMinJS(mNDVI,mNDSSI)) %>% 
  ungroup() %>% 
  filter(month == 1) %>% 
  pull(JSmin1)

JS$JSmin2 <- DeltaMeansASI %>% 
  group_by(Delta) %>% 
  mutate(JSmin2 = calcMinJS(mNDVI,mNSMI)) %>% 
  ungroup() %>% 
  filter(month == 1) %>% 
  pull(JSmin2)

#Calculate ASI and add latitude
DeltaASI <- JS %>% 
  mutate(ASI1 = sqrt(JSobs1 - JSmin1),
         ASI2 = sqrt(JSobs2 - JSmin2)) %>% #ASI1 is with NDSSI, ASI2 is with NSMI
  merge(.,DeltaDatawLocations,by="Delta") %>% 
  select(Delta,
         JSobs1,JSmin1,ASI1,
         JSobs2,JSmin2,ASI2,
         Absolute_Latitude,Lat,Lon) %>% 
  rename(AbsLat = Absolute_Latitude)
```

```{r ASI-plots}
ASIplot1 <- ggplot(DeltaASI, aes(y = AbsLat, x = ASI1)) + geom_point()
ASIplot2 <- ggplot(DeltaASI, aes(y = AbsLat, x = ASI2)) + geom_point()
ASIplot1
ASIplot2

fit1 = lm(DeltaASI$AbsLat ~ DeltaASI$ASI1)
fit2 = lm(DeltaASI$AbsLat ~ DeltaASI$ASI2)
```
```{r saveData}
save(DeltaMaxMin,DeltaPlotter,DeltaLocations,DeltaDatawLocations,DeltasCleaner,file = "AllAnalysisData.RData")
```

